### 2024-11-24 proposal.tex
Initial Overleaf Import

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2024

\PassOptionsToPackage{numbers}{natbib}

% ready for submission
\usepackage[preprint]{neurips_2024}

% to compile a preprint version, e.g., for submission to arXiv, add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2024}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2024}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2024}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{graphicx}
\usepackage{placeins}
\usepackage{calc}
\usepackage{algorithm}
\usepackage{algpseudocode}

\title{S\&DS 431 Final Project Proposal}

\author{%
  Calder K. Katyal\\
  Undergraduate\\
  Yale University\\
  New Haven, CT 06511 \\
  \texttt{calder.katyal.edu} \\
}

\begin{document}
\maketitle

\section{Original Research Goal}

\section{Dataset}
% Leave blank for now

\section{Description of Related Works}
% Leave blank for now

\section{Proposed Approaches}
% Leave blank for now

\section{Research Questions and Hypothesis}
% Leave blank for now

\section{Evaluation Metrics and Expected Results}
% Leave blank for now

\FloatBarrier

\newpage

\small
\bibliographystyle{plainnat}
\bibliography{references}

\end{document}