### 2024-11-24 proposal.tex
Begin writing

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2024

\PassOptionsToPackage{numbers}{natbib}

% ready for submission
\usepackage[preprint]{neurips_2024}

% to compile a preprint version, e.g., for submission to arXiv, add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2024}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2024}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2024}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{graphicx}
\usepackage{placeins}
\usepackage{calc}
\usepackage{algorithm}
\usepackage{algpseudocode}

\title{S\&DS 431 Final Project Proposal}

\author{%
  Calder K. Katyal; Sebastien Lowy\\
  Undergraduates\\
  Yale University\\
  New Haven, CT 06511 \\
  \texttt{calder.katyal.edu; sebastien.lowy@yale.edu} \\
}

\begin{document}
\maketitle

\section{Original Research Goal}

In this research project we propose a novel method to significantly reduce the computational cost of performing Djikstra's Shortest Path Algorithm in temporal graph data with dynamically evolving edge weights. In particular, we employ a Graph Neural Network (GNN) to predict edge importance scores and prune the graph at each timestep, significantly reducing the search space of Djikstra's algorithm. By incorporating node and edge level features as well as structural information into our model at each timestep, we can optimize Djikstra to real-world scenarios where graphs exhibit structured evolution.

\section{Dataset}
% Leave blank for now

Possibility: https://www.microsoft.com/en-us/research/publication/t-drive-trajectory-data-sample/

\section{Description of Related Works}
% Leave blank for now

\section{Proposed Approaches}
% Leave blank for now

\section{Research Questions and Hypothesis}
% Leave blank for now

\section{Evaluation Metrics and Expected Results}
% Leave blank for now

\FloatBarrier

\newpage

\small
\bibliographystyle{plainnat}
\bibliography{references}

\end{document}